Data Specialization for a Staged Lambda Calculus

### 2015-02-28 paper/relatedwork.tex

\paragraph{Frequency-reduction via pass separation.}
Perhaps the closest work to ours is the algorithm for ``data
specialization'' given by Knoblock and Ruf~\cite{knoblock96}.  Like
our work, Knoblock and Ruf are concerned about the problem of
statically splitting an explicitly staged program into two stages.
However, they only consider a simple first-order language and
straight-line programs. Their work also treats all computations
guarded by second-stage conditionals as second-stage computations,
which would prevent optimization of programs such as quickselect.

Automatic pass separation has also been a topic of interest in
programming systems for computer graphics, where, to achieve
high-performance, programs are typically manually separated into
components by frequency of execution (Separation into a pre-determined
set of per-frequency modules is \emph{required} by the execution model
of modern graphics architectures.)  The software engineering
challenges modifying multiple per-stage programs have led to
suggestions of writing graphics programs in an explicitly staged
programming language\,\cite{Proudfoot:2001,Foley:2011,He:2014}, and
deferring the task of pass separation to the compiler. However, all
prior efforts, like that of Knock and Ruf\,\cite{knoblock96} have been
limited to simple, imperative languages. We are not aware of prior
work the presents a splitting algorithm in the context of higher-order
functional languages.

>>>>

\paragraph{Frequency-reduction via pass separation.}
Perhaps the closest work to ours is the algorithm for ``data
specialization'' given by Knoblock and Ruf~\cite{knoblock96}, which
also seeks to statically split an explicitly staged program into two
stages.  However, they only consider a simple first-order language and
straight-line programs. Their work also treats all computations
guarded by second-stage conditionals as second-stage computations,
which would prevent optimization (via splitting) of programs such as
quickselect.

Automatic pass separation has also been a topic of interest in
programming systems for computer graphics, where, to achieve
high-performance, programs are typically manually separated into
components by frequency of execution (Separation into a pre-determined
set of per-frequency modules is \emph{required} by the execution model
of modern graphics architectures.)  The software engineering
challenges modifying multiple per-stage programs have led to
suggestions of writing graphics programs in an explicitly staged
programming language\,\cite{Proudfoot:2001,Foley:2011,He:2014}, and
deferring the task of pass separation to the compiler. However, all
prior efforts, like that of Knoblock and Ruf\,\cite{knoblock96} have
been limited to simple, imperative languages. We are not aware of
prior work the presents a splitting algorithm in the context of
higher-order functional languages.
